---
title: "Olympic Sports and Medals(1896-2014)"
output: html_notebook
---

More than 35,000 medals have been awarded at the Olympics since 1896. Let's Explore the data.

```{r, message=FALSE, warning=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
summer <- read.csv('summer.csv')
winter <- read.csv('winter.csv')
dictionary <- read.csv('dictionary.csv', stringsAsFactors = F)
```


While exploring the data I noticed that the country names for many country codes were missing. So, I created a separate file (missing_codes.csv) containing the missing countries (Source: https://en.wikipedia.org/wiki/List_of_IOC_country_codes). I have taken a few liberties while preparing the data to make my work easier; for example, Soviet Union has been added as 'Russia' even though it contained many countries that later became independent.

Combining the missing data and dictionary.

```{r}
missing.country <- read.csv('missing_codes.csv', stringsAsFactors = F)
country.data <- rbind(select(dictionary, Code, Country), 
                      select(missing.country, Code, Country))
```


# Summer Olympics

First lets check the Summer Olympics Data.


### Cleaning and preparing the data

Replacing the country codes in summer with the country names

```{r, warning=FALSE}
s.olympics <- tbl_df(summer) %>%
    rename(Code = Country) %>%
    left_join(., country.data, by = 'Code')

colSums(is.na(s.olympics))

```

Removing redundant values in data

```{r}
s.olympics$Discipline <- plyr::mapvalues(s.olympics$Discipline, 
                c('Artistic G.', 'Wrestling Gre-R', 'Croquet',
                  'Wrestling Free.', 'Water polo', 'Water Motorspor',
                  'Modern Pentath.', 'Canoe / Kayak F', 'BMX',
                  'Canoe / Kayak S', 'Synchronized S.', 'Rhythmic G.',
                  'Beach volley.'), 
                c('Gymnastics Artistic', 'Wrestling Greco-Roman', 'Roque',
                  'Wrestling Freestyle', 'Water Polo', 'Water Motorsport',
                  'Modern Pentathlon', 'Canoe Sprint', 'Cycling BMX',
                  'Canoe Slalom', 'Synchronized Swimming', 'Gymnastics Rhythmic',
                  'Beach Volleyball'))
```


### Top 10 countries by medals in Summer Olympics
Lets find out the top 10 countries based on number of medals won by the country.

```{r, fig.height=5, fig.width=9}
top10.countries <- s.olympics %>% 
    group_by(Country) %>%
    mutate(medal.count = n()) %>%
    select(Country, medal.count) %>%
    distinct(Country, medal.count) %>%
    arrange(desc(medal.count)) %>%
    head(10)

# top10.countries

ggplot(data = top10.countries, 
       aes(x = reorder(Country, -medal.count), y = medal.count)) +
    geom_bar(stat = 'identity', width = 0.5, fill = 'green4') +
    geom_label(aes(label = medal.count), nudge_y = 500) +
    labs(title = 'Top 10 Countries - Medal Count in Summer Olympics',
         subtitle = '1986 - 2014') +
    xlab('Country') +
    ylab('Number of Medals') +
    theme(axis.text.x = element_text(angle = 65, vjust = 0.7))
```


### Greatest Olympians of All Time (Summer Olympics)
Lets find out the top 20 athletes in Summer Olympics history based on number of individual medals won.

```{r, fig.height=5, fig.width=9}
top20.athletes <- tbl_df(s.olympics) %>%
    group_by(Athlete) %>%
    mutate(medal.count = n()) %>%
    select(Athlete, medal.count, Discipline) %>%
    distinct(Athlete, medal.count, Discipline) %>%
    arrange(desc(medal.count)) %>%
    head(20)

# top20.athletes

ggplot(top20.athletes, 
       mapping = aes(x = reorder(Athlete, medal.count), 
                     y = medal.count, 
                     fill = Discipline)) +
    geom_bar(stat = 'identity', width = 0.8) +
    geom_label(aes(label = medal.count), nudge_y = 1) +
    coord_flip() +
    labs(title = 'Greatest Olympians of All Time', 
         subtitle = 'Summer Olympics') +
    ylab('Number of Individual Medals') +
    xlab('Athlete') 
    
```

### Most Dominant Champions
Lets find the top 10 most dominant Champions (Gold Medalists) in Summer Olympics history.

### Most Profitable Sports
Lets find the top 3 sports which have produced the highest number of medals.

