---
title: "Titanic - Survival Prediction"
author: "Paresh Pradhan"
date: "5 July 2017"
output: html_document
---
## Introduction
We need to predict whether the passengers in test set surivived or perished. We will use different models and see which works best.

```{r start, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(caret)
library(stringr)

train.data <- read.csv('train.csv', stringsAsFactors = F)
test.data <- read.csv('test.csv', stringsAsFactors = F)

```

## Data Exploration
Let's explore the train dataset:

```{r EDA1}
str(train.data)
```

Lets convert Survived, Pclass and Sex into factors

```{r EDA2}
train.data$Survived <- as.factor(train.data$Survived)
train.data$Pclass <- as.factor(train.data$Pclass)
train.data$Sex <- as.factor(train.data$Sex)

summary(train.data)
```

There are 177 NA values in Age. To remove these we could just replace the NAs with the average age. But there is a better predictor of age :- The title useed in the passengers' names.

Lets extract Title from Name:

```{r EDA3}
train.data$Title <- sapply(str_split(
    string = train.data$Name, pattern = '[,.]'), '[[', 2)
train.data$Title <- as.factor(train.data$Title)
```

Lets figure out the average age for each title and then use that to replace the NAs.

