---
title: "R Notebook"
output: 
    html_notebook:
        output: readable
        author: Paresh Pradhan
---
##Introduction

Find why the best and most experienced employees are leaving prematurely.

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(ggcorrplot)
library(caret)
```


##Input and Preprocess

Reading data:
```{r}
hra.data <- read.csv('../Input/HR_comma_sep.csv', stringsAsFactors = F)
```

Modifying Data:
```{r}
hra.data <- tbl_df(hra.data)
hra.data$department <- hra.data$sales
hra.data$sales <- NULL
```

Adding Employee Status
```{r}
hra.data$emp.status <- ifelse(hra.data$left == 1, 'Left', 'Retained')
```

Adding Salary as number
```{r}
hra.data$salary.num <- hra.data$salary
hra.data$salary.num[hra.data$salary == 'low'] <- 1
hra.data$salary.num[hra.data$salary == 'medium'] <- 2
hra.data$salary.num[hra.data$salary == 'high'] <- 3
hra.data$salary.num <- as.integer(hra.data$salary.num)
```


Converting to Factors
```{r}
hra.data.fact <- hra.data
hra.data.fact[,c(-1, -2, -4)] <- lapply(hra.data.fact[,c(-1, -2, -4)], factor)
```


##Exploration

```{r}
glimpse(hra.data.fact)
```

```{r}
summary(hra.data.fact)
```



<br><br><br><html>
###Correlation

First lets check the correlation between features.

```{r, fig.height=6, fig.width=6}
g.corr <- hra.data %>% 
    select(-c(9:11)) %>% cor() %>%
    ggcorrplot(corr = ., method = 'circle', 
                     type = 'full', lab = T, lab_size = 3)

g.corr
```

Whether employees leave or stay is mostly decided by their satisfaction level.


<br><br><br><html>
###Employee Status

Distribution of Employees Retained vs Left

```{r}
g1 <- hra.data.fact %>% 
    ggplot(mapping = aes(x = emp.status)) +
    geom_bar() +
    labs(x = 'Employee Status',
         y = 'Count',
         title = 'Retained vs Left')
g1
```

Almost a third of the employees have left the company compared to those retained.



<br><br><br><html>
###Satisfaction Level

Lets compare the satisfaction levels:

```{r}
g1.1 <- hra.data %>%
    ggplot(mapping = aes(x = satisfaction_level, fill = emp.status)) +
    geom_density(alpha = 0.5) +
    labs(x = 'Satisfaction Level', 
         y = 'Count',
         fill = 'Employee Status',
         title = 'Satisfaction Level across Employee Status')
g1.1
```

Clearly, the employees that left had low satisfaction levels. But it seems there were some employees with high satisfaction level among those that left.

Lets dig deeper.



<br><br><br><html>

Lets plot satisfation vs last evaluation

```{r}
g1.2 <- hra.data.fact %>% 
    ggplot(mapping = aes(x = satisfaction_level, 
                         y = last_evaluation,
                         color = time_spend_company)) +
    geom_point(alpha = 0.5) +
    geom_abline(slope = 0, intercept = mean(hra.data$last_evaluation)) +
    geom_vline(xintercept = mean(hra.data$satisfaction_level)) +
    facet_wrap(~emp.status)
g1.2
```

There are 3 separate clusters correspoding to the 3 peaks in the previous graph for LEFT. 

In the Left plot, there is group that had both high satisfaction and high evaluation, yet they still left.

Lets dig further.



<br><br><br><html>

Plotting satisfation vs no. of projects and Average monthly hours

```{r}
g1.3 <- hra.data.fact %>%
    ggplot(mapping = aes(x = satisfaction_level,
                         y = average_montly_hours,
                         color = number_project)) +
    geom_point(alpha = 0.5) +
    geom_abline(slope = 0, intercept = mean(hra.data$average_montly_hours)) +
    geom_vline(xintercept = mean(hra.data$satisfaction_level)) +
    facet_wrap(~emp.status)
g1.3
```


Again, we get similar clusters. For the Employees LEFT plot:

Cluster 1: High Evaluation BUT too much work(High monthly hours + High no of projects) -> Low Satifaction -> Left  
Cluster 2: Low Evaluation AND too less work(Low monthly hours + Low no of projects) -> Below Average Satifaction -> Left  
Cluster 3: High Evaluation AND good amount of work(Above average monthly hours + Average no of projects) -> Above Average Satifaction -> Why Left?

Cluster-1 & 3 have employees having 4-6 years of experience.

Cluster-1 & 3 are the group of employees that the company should have retained. 

We know Cluster-1 left because of too much work. 

Why did Cluster-3 employees leave?



<br><br><br><html>

Plotting satisfation vs salary vs emp.status

```{r}
g1.4 <- hra.data.fact %>% 
    ggplot(mapping = aes(x = satisfaction_level,
                         y = last_evaluation,
                         color = time_spend_company)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~emp.status + salary)
g1.4
```


<br><br><br><html>

Plotting satisfation vs promotion vs emp.status

```{r}
g1.5 <- hra.data.fact %>% 
    ggplot(mapping = aes(x = satisfaction_level,
                         y = last_evaluation,
                         color = time_spend_company)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~emp.status + promotion_last_5years)
g1.5
```


<br><br><br><html>

Plotting satisfation vs work_accident vs emp.status

```{r}
g1.6 <- hra.data.fact %>% 
    ggplot(mapping = aes(x = satisfaction_level,
                         y = last_evaluation,
                         color = time_spend_company)) +
    geom_point(alpha = 0.5) +
    facet_wrap(~emp.status + Work_accident)
g1.6
```

Here we can see that among the employees that left, hardly anyone had high salary and almost nobody got promoted in last 5 years, especially in Cluster 1 & 3. 

Also, all three clusters in LEFT had employees who had a work accident.

That may be why they left.



<br><br><br><html>
###Correlation - Best Employees

Lets again check the correlation between features only for the best employees.

```{r, fig.height=6, fig.width=6}
g.corr2 <- hra.data %>% select(-c(9:11)) %>%
    filter(last_evaluation >= 0.75, 
           average_montly_hours >= 210, 
           time_spend_company >= 4) %>%
    cor() %>%
    ggcorrplot(corr = ., method = 'circle', 
                     type = 'full', lab = T, lab_size = 3)

g.corr2
```

The best employees that left either had low satisfaction level, too many projects and high monthly hours or had low salary didn't get promoted in last 5 years. Some might have left because of a work accident.


##Modeling

Lets use the existing data to predict which employees might leave next. We will only use the best employees for our model.

###Best Employees

```{r}
hra.data.best <- hra.data %>%
    filter(last_evaluation >= 0.75, 
           average_montly_hours >= 210, 
           time_spend_company >= 4)

hra.data.best[c(9:11)] <- lapply(hra.data.best[c(9:11)], factor)

glimpse(hra.data.best)
```

###Feature List

```{r}
feature.list <- c(1:6, 8, 10, 12)
names(hra.data.best[feature.list])
```



###Control Parameters

```{r}
tr.control <- trainControl(method = 'cv', number = 5, verboseIter = T, 
                           classProbs = T, allowParallel = T)
```


###Model Fitting

Naive bayes:

```{r}
fit.nb <- train(x = hra.data.best[feature.list], 
                y = hra.data.best$emp.status, 
                method = 'nb', metric = 'Kappa',
                trControl = tr.control)
```


Decsion Trees:

```{r}
fit.rpart <- train(x = hra.data.best[feature.list], 
                y = hra.data.best$emp.status, 
                method = 'rpart', metric = 'Kappa',
                trControl = tr.control)
```


###Resamples

```{r}
summary(resamples(list(NB = fit.nb, DT = fit.rpart)))
```


###Prediction

```{r}
prediction <- predict(fit.rpart, newdata = hra.data.best)
prediction.prob <- predict(fit.rpart, newdata = hra.data.best, type = 'prob')
```


###Confusion Matrix

```{r}
confusionMatrix(data = prediction, reference = hra.data.best$emp.status)
```

###Insights

```{r}
hra.data.best <- cbind(hra.data.best, prediction.prob)

g.insight <- hra.data.best %>%
    ggplot(mapping = aes(x = Left, y = last_evaluation)) +
    geom_point()

g.insight
```


