---
title: "What's Cooking - Predicting Cuisine from Ingredients"
author: "Paresh Pradhan"
date: "5 July 2017"
output: html_document
---

## Introduction
In this project, we will try to predict the type of cuisine from the given list of ingredients.

```{r start, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)
library(tm)
library(caret)
library(rpart)
library(rpart.plot)
library(jsonlite)
library(rattle)

data.train <- fromJSON('train.json', flatten = T)
data.test <- fromJSON('test.json', flatten = T)
```

## Understanding the data
Lets check the data:

```{r explore}
ggplot(data.train, mapping = aes(x = cuisine)) +
    geom_bar() + 
    coord_flip()
```

As you can see, Italian and Mexican cuisine dominate the data. Lets get the baseline for our predictions.

```{r baseline}
prop.table(table(as.factor(data.train$cuisine)))
```

Italian has the highest proportion of 19.7%. This means, if we predict all cuisines as Italian, we will get an accuracy of 19.7%.

Thus 19.7% is our Prediction Baseline. Lets see if our models surpass this.

## Data Preprocessing
Creating a Corpus:

```{r corpus}
ingredients <- Corpus(VectorSource(data.train$ingredients))
ingredients.prep <- tm_map(ingredients, stripWhitespace)
ingredients.prep <- tm_map(ingredients.prep, stemDocument)

inspect(ingredients.prep[[1]])
```

Creating a Term-Document Matrix:

```{r TDM}
ingredients.dtm <- DocumentTermMatrix(x = ingredients.prep)
ingredients.dtm <- removeSparseTerms(ingredients.dtm, sparse = 0.99)
```

Converting to data frame

```{r dtm_to_df}
ingredients.df <- as.data.frame(as.matrix(ingredients.dtm))
ingredients.df$cuisine <- as.factor(data.train$cuisine)
```

## Model Creation & Prediction
We will try out different models and see which performs best.

### Decision Tree
```{r dtree, message=FALSE}
set.seed(777)

tr.ctrl <- trainControl(method = 'repeatedcv', number = 5, repeats = 3, 
                        verboseIter = T, classProbs = T, 
                        selectionFunction = 'oneSE')

model.dt <- train(cuisine ~ ., data = ingredients.df, trControl = tr.ctrl,
                  method = 'rpart', metric = 'Kappa')

```

Checking Model:

```{r dtree2}
model.dt
```


Lets check the decision tree:
```{r dtree3}
rpart.plot(model.dt$finalModel)
```

Lets make the prediction and create submission file. To make prediction we need to apply the same pre-processing steps to test dataset as train.
```{r dtree4}

#pred.dt <- predict(model.dt, newdata = data.test, type = 'raw')
```

